---
title: "A new two-parameter lifetime distribution with  bathtub shape or increasing failure rate function"
output: pdf_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using Inverse transform to generate observations from the new
distribution $$
y = 1 - e^{\lambda(1 - e^{x^\beta})}
$$ $$
e^{\lambda(1 - e^{x^\beta})} = 1 - y 
$$

Taking the log on both sides $$
\lambda(1 - e^{x^\beta}) = log(1 - y)
$$

$$
1 - e^{x^\beta} = \frac{1}{\lambda}log(1-y)
$$

Taking the log of both sides $$
log(1 - \frac{1}{\lambda}log(1-y)) = x^\beta 
$$

$$
x = log(1 - \frac{1}{\lambda}log(1-y))^{\frac{1}{\beta}}
$$

where y \~ U(0,1)

```{r random-number-generator}
library(knitr)
library(purrr)
library(dplyr)
library(DataSetsUni)

rtplf <- \(n, beta, lambda){
  u <- runif(n)
  t <- ( log( 1 - (log(1 - u) / lambda) ))^(1/beta)
  return(t)
}
```

Loglikelihood of the two-parameter lifetime distribution with bathtub
shape or increasing failure rate function $$
F(X) = 1 - e^{\lambda(1 - e^{x^\beta})}
$$ $$
S(X) = 1 - F(X) = 1 - (1 - e^{\lambda(1 - e^{x^\beta})})
$$ $$
= e^{\lambda(1 - e^{x^\beta})}
$$

$$
H(X) = -log(S(X)) = -log(e^{\lambda(1 - e^{x^\beta})}) = -\lambda(1-e^{x^\beta})
$$ $$
h(x) = \frac{\partial H(X)}{x} = \lambda \beta (x^{\beta - 1})e^{x^\beta}
$$

$$
l(\theta) = \sum_i^n[\delta_ilog(h(x_i|\theta))] - \sum H(X)
$$

```{r}
ll <- \(x, par){
  beta = par[1]
  lambda = par[2]
  delta = kidney$status
  
  loglik = sum(delta * (log(lambda) + log(beta) + (beta -1 )*log(x) + x^beta)) - sum(-(lambda *(1 - exp(x^beta))))
  return(-loglik)
}
```

```{r loglikelihood-complete-data}
#Loglikelihood function
#Assuming complete data
ll_tplf <- \(x, par){
  beta = par[1]
  lambda = par[2]
  delta = rep(1, length(x))
  
  loglik = sum(delta*(log(lambda) + log(beta) + (beta -1)*log(x) + x^beta)) +
    sum(lambda*(1 - exp(x^beta)))
  return(-loglik)
}
```

Monte Carlo Simulation for N = 5000 and n = 50, 100, 150, 200, 250

```{r monte-carlo}
estimates <- \(X){
  res = optim(par = c(0.1, 0.1),
      fn = ll_tplf,
      lower = c(1e-5, 1e-5),
      method = "L-BFGS-B",
      x = X, hessian = T)
  
  invH = solve(res$hessian)
  
  res_df = tibble(beta = res$par[1],
                  lambda = res$par[2],
                  beta_var = invH[1,1],
                  lambda_var = invH[2,2])
  return(res_df)
}
```

```{r gof}
set.seed(351)
res <- \(n, N, beta, lambda){
  df = map_dfr(1:N, ~estimates(rtplf(n,beta,lambda)))
  
  tibble(n = n,
        `Bias(beta)` = mean(df$beta) - beta,
         `APE(beta)` = mean(abs((df$beta - beta)) / beta),
         `MSE(beta)` = mean((df$beta - beta)^2),
         `CP(beta)` = mean(between(rep(beta,N), 
                                   df$beta - 1.96 *
                                     sqrt(df$beta_var), 
                                   df$beta + 1.96 * 
                                     sqrt(df$beta_var))),
        `Bias(lambda)` = mean(df$lambda) - lambda,
         `APE(lambda)` = mean(abs((df$lambda - lambda)) / lambda),
         `MSE(lambda)` = mean((df$lambda - lambda)^2),
         `CP(lambda)` = mean(between(rep(lambda,N), 
                                   df$lambda - 1.96 *
                                     sqrt(df$lambda_var), 
                                   df$lambda + 1.96 * 
                                     sqrt(df$lambda_var))))
}
```

$$
\text{For} \ \beta = 0.4; \ \lambda = 2
$$

```{r}
map_dfr(c(50, 100, 150, 200, 250), ~res(.x, 5000, 0.4, 2)) |> 
  kable()
```

$$
\text{For} \ \beta = 0.8; \ \lambda = 2
$$

```{r}
map_dfr(c(50, 100, 150, 200, 250), ~res(.x, 5000, 0.8, 2)) |> 
  kable()
```

$$
\text{For} \ \beta = 1.2; \ \lambda = 2
$$

```{r}
map_dfr(c(50, 100, 150, 200, 250), ~res(.x, 5000, 1.2, 2)) |> 
  kable()

```

$$
\text{For} \ \beta = 0.4; \ \lambda = 0.5
$$

```{r}
map_dfr(c(50, 100, 150, 200, 250), ~res(.x, 5000, 0.4, 0.5)) |> 
  kable()
```

$$
\text{For} \ \beta = 0.8; \ \lambda = 0.5
$$

```{r}
map_dfr(c(50, 100, 150, 200, 250), ~res(.x, 5000, 0.8, 0.5)) |> 
  kable()
```

$$
\text{For} \ \beta = 1.2; \ \lambda = 0.5
$$

```{r}
map_dfr(c(50, 100, 150, 200, 250), ~res(.x, 5000, 1.2, 0.5)) |> 
  kable()
```

Practical Application to the strength of Glass Fibers Data in the
DataSetsUni package

```{r}
#pdf of the new distribution
dtplf <- \(x, beta, lambda){
  hx = lambda*beta*(x^(beta-1))*exp(x^beta)
  Sx = exp(lambda*(1-exp(x^beta)))
  pdf = hx*Sx
  return(pdf)
}

#Parameter estimate
fit <- optim(par = c(0.1, 0.1),
      fn = ll_tplf,
      lower = c(1e-5, 1e-5),
      method = "L-BFGS-B",
      x = data_glassf,
      hessian = T)

hist(data_glassf, freq = F)
curve(dtplf(x, fit$par[1], fit$par[2]),
      from = min(data_glassf),
      to   = max(data_glassf),
      col = "red", lwd = 2, add = TRUE)
```

```{r}
library(survival)
library(muhaz)

fit <- survfit(Surv(time, status) ~ 1, data = kidney)
haz <- muhaz(kidney$time, kidney$status)
plot(haz, main = "Estimated Hazard Function")


ll_tplf <- \(x, par){
  beta = par[1]
  lambda = par[2]
  delta = kidney$status
  
  loglik = sum(delta*(log(lambda) + log(beta) + (beta -1)*log(x) + x^beta)) +
    sum(lambda*(1 - exp(x^beta)))
  return(-loglik)
}

optim(par = c(0.1, 0.1),
      fn = ll_tplf,
      lower = c(1e-5, 1e-5),
      method = "L-BFGS-B",
      x = kidney$time,
      hessian = T)
```
